---
title: "Installation of libeemd/Rlibeemd"
author: "Jouni Helske"
date: "15 June 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Installing `Rlibeemd` is straightforward. For Windows and OSX systems you can install the binary package from CRAN in usual way. For source packages, the `libeemd` requires Gnu Scientific Library (`GSL`), so we need to install that first if we don't already have it in our system. Luckily that is straighforward in Linux (see below). 

These notes provide detailed instructions for installing `libeemd` itself especially on Windows environment. Note that if want to install `Rlibeemd` from source on Windows environment, you can also download the appropriate GSL files from here: https://www.stats.ox.ac.uk/pub/Rtools/libs.html, and then you need to add environmental variable `LIB_GSL=<path/to/gsl>`.

# Linux

On Linux, we first install `GSL` (if not already present). For example in Ubuntu:
```{bash, eval = FALSE }
sudo apt-get install libgsl2 libgsl-dev
```

And then you can just use `install.packages("Rlibeemd_*.tar.gz", repos = NULL)` in R (with `*` replaced with the relevant version number). For installing `libeemd`, you can follow the installation instructions in `libeemd` repository: https://bitbucket.org/luukko/libeemd.

# Windows

Again, we first need to install `GSL`. Unfortunately this is not as straightforward as in Linux (except maybe in Windows 10 using Linux bash?).


First, install MSYS2: www.msys2.org. Follow the instructions on the website to install the MSYS2 and update the core packages with pacman. I addition, install tar, make and mingw-toolchain:
```{r bash, eval = FALSE}
pacman -S tar make mingw-w64-x86_64-toolchain
```

Then, go to some suitable directory and download GSL:
```{r bash, eval = FALSE}
wget ftp://ftp.gnu.org/gnu/gsl/gsl-latest.tar.gz
```

And unpack:
```{r bash, eval = FALSE}
tar zxvf gsl-latest.tar.gz
```

Go to the new directory (the actual name depends on the GSL version) and install GSL:
```{r bash, eval = FALSE}
cd gsl-2.3
./configure
make
make install
```


After the installation is finished, type
```{bash, eval = FALSE}
gsl-config
```

You should see an output like this:

```{verbatim, eval = FALSE}
Usage: gsl-config [OPTION]

Known values for OPTION are:

  --prefix		show GSL installation prefix 
  --libs		print library linking information, with cblas
  --libs-without-cblas	print library linking information, without cblas
  --cflags		print pre-processor and compiler flags
  --help		display this help and exit
  --version		output version information

An external CBLAS library can be specified using the GSL_CBLAS_LIB
environment variable. The GSL CBLAS library is used by default.
```

Download the libeemd repo from bitbucket (on the left-hand side menu there should a link Download). Or you can use clone the repository with git (after installing git with `pacman -S git`).  Either way, go to the top-level directory of `libeemd` and run
```{bash, eval = FALSE}
make
make install
```

I am not a Python user so I am not sure what you need to do then?


